Submission for Course5 Week2
============================


```{r setoptions , echo=TRUE, warning=FALSE}
activity <- read.csv("activity.csv", stringsAsFactors=FALSE,header=T,na.strings = c("NA"))
library(data.table)
setDT(activity)
```

Here is a Histogram of the Total No. of steps taken per day

``` {r  , echo=FALSE}
hist(activity[,.(steps_total_per_day = sum(steps,na.rm=T)),by=date][,steps_total_per_day], main="Histogram of Total no. of Steps per day", xlab = "Total No. of Steps per day")
```

Here is the Mean Total No. of Steps taken per day

``` {r  , echo=FALSE}
activity[,.(mean_steps_per_day = mean(steps,na.rm=T)),by=date][]
```

Here is the Median Total No. of Steps taken per day

``` {r , echo=FALSE}
activity[,.(median_steps_per_day = median(steps,na.rm=T)),by=date][]
```


Here is a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):

``` {r , echo=FALSE}
plot(unique(activity[,interval]),activity[,.(avg_no_of_steps_per_interval = mean(steps,na.rm=TRUE)),by=interval][,avg_no_of_steps_per_interval], type="l", xlab = "5 minute interval", ylab = "Average No. of Steps")
```

-------------
The 5-minute interval, on average across all the days in the dataset, is internal No. 835 and the average no. of steps is 206

``` {r , echo=FALSE}
activity[,.(avg_no_of_steps_per_interval = mean(steps,na.rm=TRUE)),by=interval] [order(-avg_no_of_steps_per_interval)] [1,]
```

-------------
Imputing missing values
-----

The total number of rows with missing values in the dataset are 2304:

``` {r , echo=FALSE}
activity[,.N] - sum(complete.cases(activity))
```

The mean no. of steps for that 5-minute interval is used for imputing below:

Aftre imputing the total no. of complete case now is:

``` {r , echo=TRUE}
avg_steps_by_interval <- activity[,.(avg_no_of_steps_per_interval = mean(steps,na.rm=TRUE)),by=interval]
activity_imputed <- merge(activity, avg_steps_by_interval, by = "interval", all.x=T)
activity_imputed_final <- activity_imputed [,.(date, interval, steps = ifelse(is.na(steps),avg_no_of_steps_per_interval,steps))]
sum(complete.cases(activity_imputed_final))
```

The histogram of the total number of steps taken each day after imputing looks like below:

``` {r , echo=FALSE}
hist(activity_imputed_final[,.(total_steps_per_day = sum(steps)),by=date][,total_steps_per_day],main="Histogram of Total No. of Steps per day post imputation",xlab="Total No. of Steps per day")
```

After imputation, the histogram looks more symmetric. Before Imputation, it was skewed to the left


After Imputation, the average no. of steps per day is shown below:

``` {r , echo=FALSE}
activity_imputed_final[,.(avg_steps_per_day = mean(steps)), by=date][]
```

median steps per day:

``` {r , echo=FALSE}
activity_imputed_final[,.(median_steps_per_day = median(steps)), by=date][]
```

---------

Below is the Comparision of Average No. of steps per day before and after Imputation:

``` {r , echo=FALSE}
mean_no_of_steps_per_day_before_imputation = activity[,.(mean_steps_per_day = mean(steps,na.rm=T)),by=date]
mean_no_of_steps_per_day_before_imputation = mean_no_of_steps_per_day_before_imputation[, date_new := as.POSIXct(date,format="%Y-%m-%d")]
mean_no_of_steps_per_day_after_imputation = activity_imputed_final[,.(mean_steps_per_day = mean(steps)),by=date]
mean_no_of_steps_per_day_after_imputation = mean_no_of_steps_per_day_after_imputation[, date_new := as.POSIXct(date,format="%Y-%m-%d")]
mean_no_of_steps_per_day_diff = merge(mean_no_of_steps_per_day_before_imputation, mean_no_of_steps_per_day_after_imputation, by="date_new", all.x=T)
mean_no_of_steps_per_day_diff[, ':=' (date.x = NULL, date.y = NULL)][]
```

After Imputation, we now see the Avg no. of steps per day for all days, before Imputation it was not available for the days in which readings were not available.


Below is the Comparision of Median No. of steps per day before and after Imputation:

``` {r , echo=FALSE}
median_no_of_steps_per_day_before_imputation = activity[,.(median_steps_per_day = median(steps,na.rm=T)),by=date]
median_no_of_steps_per_day_before_imputation = median_no_of_steps_per_day_before_imputation[, date_new := as.POSIXct(date,format="%Y-%m-%d")]
median_no_of_steps_per_day_after_imputation = activity_imputed_final[,.(median_steps_per_day = median(steps)),by=date]
median_no_of_steps_per_day_after_imputation = median_no_of_steps_per_day_after_imputation[, date_new := as.POSIXct(date,format="%Y-%m-%d")]
median_no_of_steps_per_day_diff = merge(median_no_of_steps_per_day_before_imputation, median_no_of_steps_per_day_after_imputation, by="date_new", all.x=T)
median_no_of_steps_per_day_diff[, ':=' (date.x = NULL, date.y = NULL)][]
```

After Imputation, we now see the Median no. of steps per day for all days, before Imputation it was not available for the days in which readings were not available.

Weekdays and Weekends Comparision
---

Comparision of Average no. of steps per interval is shown in the chart below

``` {r , echo=FALSE, warning=FALSE, message=FALSE}
activity_imputed_final[, ':=' (day_of_the_week = ifelse(weekdays(as.POSIXct(date,'%Y-%m-%d')) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),"weekday", "weekend" ))]
activity_imputed_final$day_of_the_week <- as.factor(activity_imputed_final$day_of_the_week)
activity_by_day_of_the_week = activity_imputed_final[,.(mean_steps_per_interval = mean(steps)),by=.(interval,day_of_the_week)]
qplot(interval, mean_steps_per_interval, data=activity_by_day_of_the_week, facets = .~day_of_the_week) + geom_line() + ggtitle('Comparision of Average Steps per Interval for Weekdays and Weekends')
```

On Weekdays , around 8AM , maximum no. of steps is more than 225. This is the only peak in Weekdays. In Weekend, Peak is close to 170 steps in the morning and also multiple peaks close to this
through out the day.







